### 2008-09-03 doc/client/client_technical.tex
more text

\documentclass{book}
%\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
\newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
\newcommand{\Broker}{\texttt{Broker}}
\newcommand{\Submitter}{\texttt{Submitter}}
\newcommand{\JobDescription}{\texttt{JobDescription}}
\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
\newcommand{\JobController}{\texttt{JobController}}
\newcommand{\Job}{\texttt{Job}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\chapter{Implementation}
\label{sec:Implementation}
Clever text here.
\section{Generic Classes}
\subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
{\TargetRetriever} plugins in accordance with the input URLs using the ARC Loader (e.g. if an URL pointing to an ARC1 resource is given 
the TargetRetrieverARC1 is loaded).

@@

\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations. Contains the pure 
virtual method GetTargets which is to be implemented by the specialized class. While it is not mandatory it is recommended 
that the specialized class divides this function into two components: QueryIndex and InterrogateTarget. The former is 
responsible for querying index servers, and if the result yields an addresss to a different index server then the 
specialized {\TargetRetriever} should call itself recursively with the found url. Discovered computing services are sent 
to the InterrogateTarget function (or if the {\TargetRetriever} was initialized with a computing service url then it should 
jump directly to this point) which collects the detailed information about the service.

\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of running a grid job. It
serves as input to the broker which is foreseen to be able to select between different {\ExecutionTarget}s from different 
grid flavours without apriori knowing their difference. The {\ExecutionTarget} class mimics the glue2 information model (with 
a flattened structure), and thus a mapping between attributes from other information systems into the glue2 format is needed. 
Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares. All information can be printed by 
invoking the {\ExecutionTarget}'s ``Stat'' command which prints all variables (if they have been filled during the information 
retrieval process).

Following a broker decision jobs are to be submitted. Since all information about the selected computation resides within the 
selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a submitter (see next section) capable of submitting a 
job to the resource which is represents.

\subsection{{\Broker}} is not yet implemented and thus not described here.

\subsection{{\JobDescription}} The {\JobDescription} is responsible for handling all job descriptions. Thus it is capable of handling the xRSL, JDL and 
JSDL job description languages and to some extent also translate between them. Validation (syntax and semantical) is performed 
automatically when the source is set. Features easy retrieval of list of local input files which should be uploaded to the 
computing service.

\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin).  It submits job(s) to a service and 
uploads (by the job needed) local input files. Returns an XML node containing all needed information about the job 
for later job management. The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the 
{\ExecutionTarget} populates (through XML config element) the {\Submitter} with information about submission endpoint (URL) 
and job description languages ``spoken'' by the target.

\subsection{{\JobSupervisor}} is responsible for loading the appropriate {\JobController} for managing a job running on a specific 
grid flavour. In order to match a job identification string (jobid) with a grid flavour a lookup in the local xml file is 
performed. The {\JobSupervisor} main functionality is to return a list of {\JobController}s which are capable of managing all jobs 
addressed by a certain command.

\subsection{{\JobController}} is both base class for grid specific specializations, but also the implementor of all public functionality 
offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. Upon initialization the 
{\JobController} converts the information stored in the local XML to Jobs which are stored in a JobControler internal store called JobsStore. 
The public functions of the {\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations 
for issueing the command. This ensures consistency in how the commands are handled between the different grid flavours. Jobs that are 
permanently removed following a command are removed from both the JobStore and local xml file. Thus it is foreseen that the {\JobController}s 
may be used in a future graphical user interface where it becomes a persistent object continously monitoring jobs.

\subsection{{\Job}} is generic job class for storing all job related information. Attributes are derived from the glue2 information model 
and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present mapping schema.

\subsection{UserConfig}

@@

\section{Support Classes}
\subsection{URL}
\subsection{Logger}

>>>>

\documentclass{book}
%\documentclass{article}                            %for shorter notes
\usepackage{graphicx}                              %for PNG images (pdflatex)
%\usepackage{graphics}                              %for EPS images (latex)
\usepackage[linkbordercolor={1.0 1.0 0.0}]{hyperref} %for \url tag
\usepackage{color}                                 %for defining custom colors
\usepackage{framed}                                %for shaded and framed paragraphs
\usepackage{textcomp}                              %for various symbols, e.g. Registered Mark
\usepackage{geometry}                              %for defining page size
\usepackage{longtable}                             %for breaking tables
%
\geometry{verbose,a4paper,tmargin=2.5cm,bmargin=2.5cm,lmargin=2.5cm,rmargin=2cm}
\hypersetup{
  pdfauthor = {Author Name},
  pdftitle = {Paper title},
  pdfsubject = {Paper subject},
  pdfkeywords = {Paper,keyword,comma-separated},
  pdfcreator = {PDFLaTeX with hyperref package},
  pdfproducer = {PDFLaTeX}
}
%
\bibliographystyle{IEEEtran}                       %a nice bibliography style
%
\def\efill{\hfill\nopagebreak}%
\hyphenation{Nordu-Grid}
\setlength{\parindent}{0cm}
\setlength{\FrameRule}{1pt}
\setlength{\FrameSep}{8pt}
\addtolength{\parskip}{5pt}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\renewcommand{\arraystretch}{1.3}
\newcommand{\dothis}{\colorbox{shadecolor}}
\newcommand{\globus}{Globus Toolkit\textsuperscript{\textregistered}~2~}
\newcommand{\GT}{Globus Toolkit\textsuperscript{\textregistered}}
\newcommand{\ngdl}{\url{http://ftp.nordugrid.org/download}~}
\newcommand{\ACC}{\texttt{ACC}}
\newcommand{\TargetGenerator}{\texttt{TargetGenerator}}
\newcommand{\TargetRetriever}{\texttt{TargetRetriever}}
\newcommand{\ExecutionTarget}{\texttt{ExecutionTarget}}
\newcommand{\Broker}{\texttt{Broker}}
\newcommand{\Submitter}{\texttt{Submitter}}
\newcommand{\JobDescription}{\texttt{JobDescription}}
\newcommand{\JobSupervisor}{\texttt{JobSupervisor}}
\newcommand{\JobController}{\texttt{JobController}}
\newcommand{\Job}{\texttt{Job}}
\newcommand{\UserConfig}{\texttt{UserConfig}}
\newcommand{\ClientInterface}{\texttt{ClientInterface}}
\definecolor{shadecolor}{rgb}{1,1,0.6}
\definecolor{salmon}{rgb}{1,0.9,1}
\definecolor{bordeaux}{rgb}{0.75,0.,0.}
\definecolor{cyan}{rgb}{0,1,1}
%
%----- DON'T CHANGE HEADER MATTER
\begin{document}
\def\today{\number\day/\number\month/\number\year}

@@

\chapter{Implementation}
\label{sec:Implementation}
Clever text here.
\section{Generic Classes}
\subsection{{\ACC}}

\subsection{{\TargetGenerator}} The {\TargetGenerator} class is the main class for resource discovery and information retrieval. It loads 
{\TargetRetriever} plugins in accordance with the input URLs using the ARC Loader (e.g. if an URL pointing to an ARC1 resource is given 
the TargetRetrieverARC1 is loaded).

@@

\subsection{{\TargetRetriever}} The {\TargetRetriever} is base class for {\TargetRetriever} grid middleware specializations and 
inherits from the \texttt{ACC} base class in order to be loadable by the ARC \texttt{Loader}. It is designed to work on conjunction 
with the {\TargetGenerator} contains and contains the pure virtual method

\begin{shaded}
\begin{verbatim}
virtual void GetTargets(TargetGenerator& mom, int targetType,
                        int detailLevel) = 0;
\end{verbatim}
\end{shaded}

which is to be implemented by the specialized class. While it is not mandatory it is recommended that the specialized class 
divides this method into two components: QueryIndex and InterrogateTarget. The former is handles index server queries, and the 
latter the computing service queries and {\ExecutionTarget} preparation.

If an index server query yields an url to a different index server then the one queried, then the {\TargetRetriever} should 
call itself recursively creating a new {\TargetRetriever} for the discovered index server.

\subsection{{\ExecutionTarget}} The {\ExecutionTarget} is the class representation of a computing resource (queue) capable of 
executing a grid job. It serves as input to the {\Broker} which is foreseen to be able to select between different 
{\ExecutionTarget}s from different grid flavours without apriori knowing their difference. The {\ExecutionTarget} class 
mimics the glue2 information model (with a flattened structure), and thus a mapping between attributes from other information 
systems into the glue2 format is needed.  Appendix XX shows the current mapping for the production ARC, ARC1 and gLite middlewares.

All attributes of the {\ExecutionTarget} can be printed by the method

\begin{shaded}
\begin{verbatim}                                                                                                                            
void Print(bool longlist) const;
\end{verbatim}
\end{shaded}

Following a broker decision jobs are to be submitted. Since all information about the selected computing service resides within the 
selected {\ExecutionTarget}, the {\ExecutionTarget} is capable of returning a {\Submitter} capable of submitting a 
job to the service which is represents.

\begin{shaded}
\begin{verbatim}
Submitter *GetSubmitter() const;
\end{verbatim}
\end{shaded}

\subsection{{\Broker}} The {\Broker} is not yet implemented and thus not described here.

\subsection{{\JobDescription}} When addressing interoperability it is of paramount importance to transparently address grid job 
descriptions written in different job description languages by translating them automatically. In the ARClib library this functionality 
is implemented in the {\JobDescription} class. This is a generic class that takes a job description (string )as input

\begin{shaded}
\begin{verbatim}
setSource( const std::string source ) throw(JobDescriptionError);
\end{verbatim}
\end{shaded}

in any supported formats (currently xRSL, JDL, JSDL), converts and stores it in a JSDL-like internal job description format. The identification 
of the source's description format is internally handled by a \texttt{JobDescriptionOrderer} class which identifies the format by pattern
matching.

Different operations, like getting the job description in other formats or getting job-related information, can be performed using the 
description-independent functions of this class, e.g.:

\begin{shaded}
\begin{verbatim}
getProduct(std::string& product, std::string format="JSDL") throw(JobDescriptionError);
\end{verbatim}
\end{shaded}

The {\JobDescription} class has tree back-end classes, sometimes referred to as back-end or translator modules, corresponding to the three 
supported job description languages. The {\JobDescription} class chooses the appropriate back-end module according to the pattern match 
performed by the \texttt{JobDescriptionOrderer}, and uses the back-end module for parsing and generating the job descriptions.

\subsection{{\Submitter}} {\Submitter} is base class for grid specific specializations (plugin). It submits job(s) to the service 
it represents and uploads (by the job needed) local input files.

\begin{shaded}
\begin{verbatim}
virtual bool Submit(JobDescription& jobdesc, XMLNode& info) = 0;
\end{verbatim}
\end{shaded}

The \texttt{Submit} method fills the XML node \texttt{info} with all needed information about the job for later job management. 
The {\Submitter} is returned by the {\ExecutionTarget} selected for job execution and thus the {\ExecutionTarget} populates 
(through XML config element) the {\Submitter} with information about submission endpoint (URL) and job description languages 
``spoken'' by the target.

\subsection{{\JobSupervisor}} The {\JobSupervisor} is responsible for loading the appropriate {\JobController}s for managing 
jobs running on a certain grid flavour. Job manipulation can be performed either on individual jobs or on groups of jobs (e.g. 
all jobs running on certain cluster or all jobs with job state ``FINISHED''), and in order to translate the by user given 
information into a set of loadable {\JobController}s the {\JobSupervisor} makes extensive use of the local XML file housing 
information about all active jobs. Thus the {\JobSupervisor} constructor becomes

\begin{shaded}
\begin{verbatim}
JobSupervisor(const std::list<std::string>& jobs,
              const std::list<std::string>& clusterselect,
              const std::list<std::string>& clusterreject,
              const std::string joblist);
\end{verbatim}
\end{shaded}

where \texttt{jobs} is a list of job identification strings (jobids), \texttt{clusterselect} is a list of selected computing services, 
\texttt{clusterreject} is a list of rejected computing services and \texttt{joblist} is a string pointing to the local XML file 
with job information. Syntax for \texttt{clusterselect} and \texttt{clusterreject} is identical to the one used by the {\TargetGenerator}
for for internal consistency.

Although being loaded by the {\JobSupervisor} the {\JobController} objects truly resides within the ARC Loader which is a member 
of the {\JobSupervisor} class. In order to get handles on the {\JobController}s the inline method

\begin{shaded}
\begin{verbatim}
std::list<Arc::JobController*> GetJobControllers(){return JobControllers;}
\end{verbatim}
\end{shaded}

returns a list of pointers to the loaded {\JobController}s.

\subsection{{\JobController}} The {\JobController} is both base class for grid specific specializations, but also the implementor of all 
public functionality offered by the {\JobController}s. In other words all virtual functions of the {\JobController} are private. The 
initialization of a (specialized) {\JobController} object takes two steps. First the constructor has to invoked

\begin{shaded}
\begin{verbatim}
JobController(Arc::Config *cfg, std::string flavour);
\end{verbatim}
\end{shaded}

This is usually done by the ARC Loader which sees to that the {\JobController} receives information about its flavour (grid) and 
the local XML file containing information about all active jobs (flavour independent). Next step is the filling of the {\JobController}'s
job pool \texttt{JobStore} which is the pool of jobs that the {\JobController} can manage.

\begin{shaded}
\begin{verbatim}
void FillJobStore(std::list<std::string> jobs,
                  std::list<std::string> clusterselect,
                  std::list<std::string> cluterreject);
\end{verbatim}
\end{shaded}

Here \texttt{jobs}, \texttt{clusterselect} and \texttt{clusterreject} have the same interpretation as in the {\JobSupervisor} constructor. 
However, while the latter only had to identify which {\JobController}s should be loaded, the {\JobController} itself has to decode the 
supplied lists into a list of jobs with non-overlapping entries which will be filled into the JobStore. Due to possible conflicts 
within the supplied lists, rules have to be set up for dealing with contradictions:

\begin{enumerate}
\item{If the \texttt{jobs} list has entries, fill \texttt{JobStore} with the by user requested jobs.}
\item{If the \texttt{clusterselect} list has entries, fill \texttt{JobStore} with the jobs running on the selected clusters.}
\item{If clusters are rejected and \texttt{JobStore} has entries, remove from \texttt{JobStore} all jobs running on rejected clusters.}
\item{If \texttt{jobs} and \texttt{clusterselect} are both empty lists, fill \texttt{JobStore} with all jobs except those 
running on possible rejected clusters.}
\end{enumerate}

The steps above completes the initialization of the {\JobController} which is now ready for handling jobs. The public functions of the 
{\JobController} offer to get (download), clean, cancel, etc one or more jobs and uses the private specializations for issueing the command. 
Here examplified by the \texttt{Stat} command:

\begin{shaded}
\begin{small}
\begin{verbatim}
void JobController::Stat(const std::list<std::string>& jobs,
                    const std::list<std::string>& clusterselect,
                    const std::list<std::string>& clusterreject,
                    const std::list<std::string>& status,
                    const bool longlist,
                    const int timeout){

  GetJobInformation();

  //Add here functionality for identifying jobs according to clusterselect, clusterreject & status

  for(std::list<Arc::Job>::iterator jobiter = JobStore.begin();jobiter!= JobStore.end();jobiter++){
    if(jobiter->State.empty()){
      std::cout<<Arc::IString("Job state information not found:%s",jobiter->JobID.str())<<std::endl;
      Time now;
      if(now - jobiter->LocalSubmissionTime < 90){
        std::cout<<Arc::IString("This job was very recently "
                                "submitted and might not yet" 
                                "have reached the information-system")<<std::endl;
      }
      continue;
    }
    jobiter->Print(longlist);
  }
}
\end{verbatim}
\end{small}
\end{shaded}

The \texttt{Stat} command prints the job information to screen and in order to do so the {\JobController} has to query local information
server for the latest status. Due to different protocols used for different grid flavours (e.g. ldap for production ARC), the 
\texttt{GetJobInformation} has to be grid flavour specific and is only declared as a private virtual method within the {\JobController}
base class. For details about the flavour specific implementations see Section~\ref{sec:plugins}.

It should be noted that the \texttt{Stat} example above appears to have several unused variables (arguments). This is an artifact 
caused by the ongoing development in which the {\JobController} is being prepared for usage in a future graphical user interface (GUI).
In a GUI it would be desirable to make the {\JobController} a persistent object which exists throughout the lifetime of the GUI and 
which is capable of monitoring and managing all jobs running on a certain grid flavour. In order to do so the \texttt{JobStore} should 
become a pool of all jobs, and the public functions should take arguments which identifies which jobs from the pool should be subject 
to the command. In the present implementation commands are performed on all jobs in \texttt{JobStore}.

\subsection{{\Job}} The {\Job} is a generic job class for storing all job related information. Attributes are derived from the 
glue2 information model and thus a mapping is needed for non glue2 compliant grid middlewares. Appendix XX shows the present 
mapping schema.

All attributes of the {\Job} can be printed by the method

\begin{shaded}
\begin{verbatim}                                                                                                                            
void Print(bool longlist) const;
\end{verbatim}
\end{shaded}

\subsection{{\UserConfig}} The {\UserConfig} class handles the client setup i.e. proxy, certificate and key location, user and system 
configuration and local joblist location. Upon initialization (constructor) the {\UserConfig} locates the user files

\begin{shaded}
\begin{verbatim}                                                                                                                            
$HOME/.arc/client.xml
$HOME/.arc/jobs.xml
\end{verbatim}
\end{shaded}

and if either of them is non-existing a default (empty) one is created.

The {\UserConfig} has two main public methods

\begin{shaded}
\begin{verbatim}
const std::string& JobsFile();
const XMLNode ConfTree();
\end{verbatim}
\end{shaded}

where \texttt{JobsFile()} returns the string pointing to the jobs.xml file while \texttt{ConfTree()}
returns a configuration XML object which is the merge between the user and system configurations.
In order to do this the method has to locate the system configuration and resolve possible conflicts 
with the user configuration. This proceeds through the following chain of actions:

\begin{enumerate}
\item{Try reading system configuration from \texttt{<ARC Install Location>/etc/arcclient.xml}}
\item{If the previous step failed try reading system configuration from \texttt{/etc/arcclient.xml}}
\item{Merge system and user configuration by adding all system configuration not already listed 
in the user configuration to the latter.}
\item{Check if \texttt{\$X509\_USER\_PROXY} is set, if set add this as \texttt{ProxyPath} to configuration}
\end{enumerate}

\subsection{{\ClientInterface}}

@@

\section{Support Classes}
\subsection{URL}
\subsection{Logger}
\subsection{XMLNode}
